Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-24 paper/main.tex
standardized phrasing of lecture 1 title

While the time courses in Figure~\ref{fig:knowledge-timeseries}A and C provide
detailed \textit{estimates} about participants' knowlege, these estimates are
of course only \textit{useful} to the extent that they accurately reflect what
participants actually know. As one sanity check, we anticipated that the
knowledge estimates should reflect a content-specific ``boost'' in
participants' knowledge after watching each lecture. In other words, if
participants learn about each lecture's content when they watch each lecture,
the knowledge estimates should capture that. After watching the \textit{Four
Fundamental Forces} lecture, participants should exhibit more knowledge for the
content of that lecture than they had before, and that knowledge should persist
for the remainder of the experiment. Specifically, knowledge about that
lecture's content should be relatively low when estimated using Quiz~1
responses, but should increase when estimated using Quiz~2 or 3 responses
(Fig.~\ref{fig:knowledge-timeseries}B). Indeed, we found that participants'
estimated knowledge about the content of the \textit{Four Fundamental Forces}
was substantially higher on Quiz~2 versus Quiz~1 ($t(49) = 8.764,~p < 0.001$)
and on Quiz~3 versus Quiz~1 ($t(49) = 10.519,~p < 0.001$). We found no reliable
differences in estimated knowledge about that lecture's content on Quiz~2
versus 3 ($t(49) = 0.160,~p = 0.874$). Similarly, we hypothesized (and
subsequently confirmed) that participants should show greater estimated
knowledge about the content of the \textit{Birth of Stars} lecture after
(versus before) watching it (Fig.~\ref{fig:knowledge-timeseries}D).
Specifically, since participants watched that lecture after taking Quiz~2 (but
before Quiz~3), we hypothesized that their knowledge estimates should be
relatively low on Quizzes~1 and 2, but should show a ``boost'' on Quiz~3.
Consistent with this prediction, we found no reliable differences in estimated
knowledge about the \textit{Birth of Stars} lecture content on Quizzes~1 versus
2 ($t(49) = 1.013,~p = 0.316$), but the estimated knowledge was substantially
higher on Quiz~3 versus 2 ($t(49) = 10.561,~p < 0.001$) and Quiz~3 versus 1
($t(49) = 8.969,~p < 0.001$).

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see
    \nameref{subsec:knowledge-maps}). The topic trajectories of the two
    lectures are indicated by dotted lines (blue: Lecture 1; green: Lecture 2),
    and the coordinates of each question are indicated by dots (light blue:
    Lecture 1-related; light green: Lecture 2-related; purple: general physics
    knowledge). Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps
    follow the same general format as the knowledge maps in Panel A, but here
    the shading at each coordinate indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' between the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the weighted blend of words underlying the topic proportions represented at
    the corresponding colored star's location on the maps. In each word cloud,
    the words' relative sizes correspond to their relative weights at the
    starred location, and their colors indicate their relative weights in the
    \textit{Four Fundamental Forces} (blue) versus \textit{Birth of Stars}
    (green) lectures, on average, across all timepoints' topic vectors.}

>>>>

While the time courses in Figure~\ref{fig:knowledge-timeseries}A and C provide
detailed \textit{estimates} about participants' knowlege, these estimates are
of course only \textit{useful} to the extent that they accurately reflect what
participants actually know. As one sanity check, we anticipated that the
knowledge estimates should reflect a content-specific ``boost'' in
participants' knowledge after watching each lecture. In other words, if
participants learn about each lecture's content when they watch each lecture,
the knowledge estimates should capture that. After watching the \textit{Four
Fundamental Forces} lecture, participants should exhibit more knowledge for the
content of that lecture than they had before, and that knowledge should persist
for the remainder of the experiment. Specifically, knowledge about that
lecture's content should be relatively low when estimated using Quiz~1
responses, but should increase when estimated using Quiz~2 or 3 responses
(Fig.~\ref{fig:knowledge-timeseries}B). Indeed, we found that participants'
estimated knowledge about the content of \textit{Four Fundamental Forces}
was substantially higher on Quiz~2 versus Quiz~1 ($t(49) = 8.764,~p < 0.001$)
and on Quiz~3 versus Quiz~1 ($t(49) = 10.519,~p < 0.001$). We found no reliable
differences in estimated knowledge about that lecture's content on Quiz~2
versus 3 ($t(49) = 0.160,~p = 0.874$). Similarly, we hypothesized (and
subsequently confirmed) that participants should show greater estimated
knowledge about the content of the \textit{Birth of Stars} lecture after
(versus before) watching it (Fig.~\ref{fig:knowledge-timeseries}D).
Specifically, since participants watched that lecture after taking Quiz~2 (but
before Quiz~3), we hypothesized that their knowledge estimates should be
relatively low on Quizzes~1 and 2, but should show a ``boost'' on Quiz~3.
Consistent with this prediction, we found no reliable differences in estimated
knowledge about the \textit{Birth of Stars} lecture content on Quizzes~1 versus
2 ($t(49) = 1.013,~p = 0.316$), but the estimated knowledge was substantially
higher on Quiz~3 versus 2 ($t(49) = 10.561,~p < 0.001$) and Quiz~3 versus 1
($t(49) = 8.969,~p < 0.001$).

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see
    \nameref{subsec:knowledge-maps}). The topic trajectories of the two
    lectures are indicated by dotted lines (blue: Lecture 1; green: Lecture 2),
    and the coordinates of each question are indicated by dots (light blue:
    Lecture 1-related; light green: Lecture 2-related; purple: general physics
    knowledge). Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps
    follow the same general format as the knowledge maps in Panel A, but here
    the shading at each coordinate indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' between the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see Supplementary
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the weighted blend of words underlying the topic proportions represented at
    the corresponding colored star's location on the maps. In each word cloud,
    the words' relative sizes correspond to their relative weights at the
    starred location, and their colors indicate their relative weights in
    \textit{Four Fundamental Forces} (blue) versus \textit{Birth of Stars}
    (green) lectures, on average, across all timepoints' topic vectors.}
